\documentclass{article}
\usepackage{fullpage}
\usepackage{epsfig}
\usepackage{pdfsync}
\usepackage{amsfonts}

\begin{document}

\title{Documentation for the Multivariate Normal Gibbs sampler}
\author{Anand}
\maketitle

The Gibbs samplers in \texttt{Normal.py} can be applied to stochs $x$ whose children are distributed as follows: For each child $c_i$,
\begin{eqnarray*}
	(c_i | x, F_i, a_i, \tau_i) \sim \textup{N}(F_i x + a_i, \tau_i).
\end{eqnarray*}

\section{Conjugate}\label{conj}
The step method \texttt{cMVNormalWithMVNormalChildren} applies if $x$'s prior is normal:
\begin{eqnarray*}
	(x|\mu_p,\tau_p) \sim \textup{N}(\mu_p,\tau_p).
\end{eqnarray*}
In this case, the log conditional probability up to constants not involving $x$, $\sum_i \log p_i(c_i|x,\ldots) +\log p_x(x|\ldots)$, is equal to
\begin{eqnarray*}
	-\frac{1}{2}(c_i-F_ix-a_i)^T\tau_i(c_i-F_ix-a_i) - \frac{1}{2}(x-\mu_p)^T\tau_p(x-\mu_p),
\end{eqnarray*}
which after completing the square yields
\begin{eqnarray*}
	(x|\{c_i\},\ldots)\sim\textup{N}(\mu,\tau),\\
	\tau = \tau_p + \sum_i F_i^T \tau_i F_i,\\
	\mu = \tau^{-1}(\tau_p\mu_p + \sum_i F_i^T\tau_i(x_i-a_i)).
\end{eqnarray*}

This step method's constructor arguments are:
\begin{itemize}
	\item \texttt{stoch}: The stoch representing $x$. 
	\item \texttt{F\_dict}: A dictionary, indexed by child objects, whose values are dtrms whose values are the $F$ matrices. 
	\item \texttt{a\_dict}: Same as \texttt{F\_dict}, but the values of the dtrms are the $a$ arrays.
	\item \texttt{tau\_dict}: The obvious.
	\item \texttt{prior\_mu}: A dtrm whose value is the prior mean of $x$.
	\item \texttt{prior\_tau}: Same as \texttt{prior\_mu}, but with the precision matrix.    
\end{itemize}

\section{Nonconjugate}\label{non}
The step method \texttt{cMVNormalWithMVNormalChildren} applies if $x$'s prior $p_x(x)$ is non-normal. In this case, the log joint probability of $x$ up to unimportant constants, $\sum_i \log p_i(c_i|x,\ldots) + \log p_x(x)$, is equal to
\begin{eqnarray*}
	-\sum_i\frac{1}{2}(c_i-F_ix-a_i)^T\tau_i(c_i-F_ix-a_i) +\log p_x(x).
\end{eqnarray*}
A Metropolis-Hastings step algorithm for this distribution which tends to be efficient is the following:
\begin{enumerate}
	\item Propose a value $x^p$ for $x$ as if it had an uninformative prior ($\tau_p = \epsilon I$, with $\epsilon << 1$). That is,
	\begin{eqnarray*}
		(x_p|\{c_i\},\ldots)\sim\textup{N}(\mu,\tau),\\
		\tau = \sum_i F_i^T \tau_i F_i,\\
		\mu = \tau^{-1}\sum_i F_i^T\tau_i(x_i-a_i).
	\end{eqnarray*}	
	\item Accept the jump with probability
	\begin{eqnarray*}
		\min\left\{1,\frac{p_x(x^p)}{p_x(x)}\right\}.
	\end{eqnarray*}
\end{enumerate}

This step method's constructor arguments are simply:
\begin{itemize}
	\item \texttt{stoch}: The stoch representing $x$. 
	\item \texttt{F\_dict}: A dictionary, indexed by child objects, whose values are dtrms whose values are the $F$ matrices. 
	\item \texttt{a\_dict}: Same as \texttt{F\_dict}, but the values of the dtrms are the $a$ arrays.
	\item \texttt{tau\_dict}: The obvious.
\end{itemize}

\section{Things that would be nice for Gibbs sampling in general}
I started a class called Gibbs in \texttt{pymc/special\_StepMethods/GibbsSampler.py}, but couldn't figure out how to make it work well. It would be nice if:  
\begin{itemize}
	\item We come up with some consistent class factory for making conjugate/ nonconjugate pairs of Gibbs samplers, maybe similar to what we're doing in distributions.py.	
	\item The accessory dictionaries that get passed in to the constructor should be allowed to contain stochs, dtrms, or simple ndarrays. The nodes should have their values extracted automatically somehow.
	\item Are we willing to insist on consistent parent keying schemes, to avoid having to pass in the accessory dictionaries? If so, we need to standardize across all Gibbs samplers and distributions.py. Compatibility for weird models will require heavy use of utility dtrms.
	\item Each Gibbs sampler holds a OneAtATimeMetropolis instance called `default' or something which it uses if a Gibbs step fails.
\end{itemize} 
\end{document}